---
title: "Homework #0: Hello DS-6030"
---


# Getting Help

The purpose of this (ungraded) homework is to help get you prepared for the semester. Don't panic if you don't immediately know the answers to some of these. I expect everyone will need to look things up. Take note of the areas that are rusty and plan to spend a bit of extra time to get up to speed. If some of these items are not even remotely familiar to you, then you probably have not satisfied the pre-requisite material; review the course syllabus and speak to me about any questions.

The teaching staff (TA and myself) are here to help! Don't wait too long before asking for help and do let us know right away if you are starting to fall behind. I will also publish the solutions after due date. It is **highly encouraged** that you study the posted solutions.

# Course Tools

## a. Working with Quarto


1. Install or update quarto <https://quarto.org/docs/get-started/> to version 1.8.26 (or greater).
    - If you are using VSCode, install the [Quarto extension](https://marketplace.visualstudio.com/items?itemName=quarto.quarto>).

2. Course Homework Style

In order to help with grading, you need to use special formatting to generate the homework html file. 

Download the following files and put them in the top level of where you will keep your homework. 

An example directory structure is:
```
hw/
├─ _quarto.yml
├─ course-theme.css
├─ hw0.qmd
├─ hw0.html
├─ hw1.qmd
├─ hw1.html
...
```


## b. Update IDE

Update your IDE and (optionally) try a new one! These will all work great for this course. 

- [Positron](https://positron.posit.co/) (2026.01.0-147)

- [RStudio](https://posit.co/download/rstudio-desktop/) (2026.01.0+392) 

- [VSCode](https://code.visualstudio.com/) (1.108)
    - If using VSCode, install the [Quarto extension](https://marketplace.visualstudio.com/items?itemName=quarto.quarto>)


## c. Update R 

- Use the latest version (R 4.5.2, Quarto 1.8.26).
    - Installation help R: <https://cran.rstudio.com/>
    
- Install/Update the following packages we will meet during this course:
  - Dynamic report generation: `rmarkdown`, `knitr`, `quarto`
  - Utility: `remotes`
  - Working with Data: `tidyverse`
  - Data: `ISLR`, `moderndive`, `MASS`
  - Resampling: `boot`, `rsample`
  - Modeling: `tidymodels`
  - Regression: `glmnet`, `FNN`,
  - Classification: `e1071`,
  - Trees: `rpart`, `rpart.plot`, `randomForest`, `ranger`
  - Ensembles: `xgboost`, `lightgbm`, `bonsai`
  <!-- - Density Estimation: `fitdistrplus`, `ks` -->
  <!-- - Clustering: `mclust`, `mixtools`, `dbscan` -->
  <!-- - Association Analysis: `arules` -->
  <!-- - Networks: `igraph` -->
  - Forecasting: `fpp3`

You can see which of your existing packages need updating by running:
```{r, echo=TRUE, eval=FALSE}
old.packages() %>% as_tibble()
```
Or use the `update.packages()` function to update them. 

Note: Do not call `install.packages()` in this Quarto document; it only needs to be done once from the console. However you will need to use `library()` in Quarto since it needs to be called every time the document is compiled.

Below are some resources for help with modern R.

::: {.callout-tip collapse="true" title = "R Resources"}

### Tidyverse Resources

Read the following sections in [R for Data Science 2e](https://r4ds.hadley.nz/):

- The Whole Game
    - Chapters 2-9

- Visualize
    - Chapters 10-12

- Transform
    - 12-15, 19

- Program
    - 26-27

- Communicate
    - 29


### Cheatsheets and resources

[Posit Cheatsheets](https://posit.co/resources/cheatsheets/)

### **RStudio and Quarto**

- [RStudio IDE Cheatsheet](https://rstudio.github.io/cheatsheets/html/rstudio-ide.html)

- [Quarto Website](https://quarto.org/docs/authoring/markdown-basics.html)

- [Quarto Cheatsheet](https://rstudio.github.io/cheatsheets/html/quarto.html)

- [Latex Cheatsheet](https://wch.github.io/latexsheet/latexsheet-0.png)

### **R**

- [Base R](https://rstudio.github.io/cheatsheets/base-r.pdf)

- [Data Visualization Cheatsheet](https://rstudio.github.io/cheatsheets/html/data-visualization.html)
    - [`ggplot2` website](https://ggplot2.tidyverse.org/)

- [Tidy Data Cheatsheet](https://rstudio.github.io/cheatsheets/html/tidyr.html)
    - [`tidyr` website](https://tidyr.tidyverse.org/)
    
- [Data Transform Cheatsheet](https://rstudio.github.io/cheatsheets/html/data-transformation.html)
    - [`dplyr` website](https://dplyr.tidyverse.org/)    

- [Factors with forcats Cheatsheet](https://rstudio.github.io/cheatsheets/html/factors.html)
    - [`forcats` website](https://forcats.tidyverse.org/)

- [Data Import Cheatsheet](https://rstudio.github.io/cheatsheets/html/data-import.html)
    - [`readr` website](https://readr.tidyverse.org/)

- [Apply Functions Cheatsheet](https://rstudio.github.io/cheatsheets/html/purrr.html)
    - [`purrr` website](https://purrr.tidyverse.org/)

### **Python with RStudio/RMarkdown** (Optional)

- [Python with R and Reticulate Cheatsheet](https://rstudio.github.io/cheatsheets/html/reticulate.html)

:::


## d. Python Package Manager 

I recommend [uv](https://docs.astral.sh/uv/) for python package management. 

- To get you started, modify and add the following `project.toml` file to your class directory:

::: {.callout-tip collapse="true" title = "pyproject.toml"}

```
[project]
name = "DS6030"
version = "0.1.0"
requires-python = ">=3.14"

dependencies = [
  # numerical computing
  "numpy",
  "scipy",

  # data manipulation
  "pandas",
  "polars",
  "ibis-framework",

  # machine learning
  "scikit-learn",
  "statsmodels",

  # plotting and visualization
  "matplotlib",
  "seaborn",
  "plotnine",

  # notebooks and execution
  "jupyterlab",
  "ipykernel"
]
```
:::

- Then run the following to add the python and packages from the project.toml:

```
uv python install
uv sync
```



# Problem 1: Math Notation

## a. Least Squares

What are the equations for the least squares coefficients in linear regression (in matrix notation)? Use $X$ for the model/design/predictor matrix, and $Y$ the vector of outcomes.

::: {.callout-note title="Solution"}
Add solution here
:::

## b. MLE

Let $x_1, x_2, \ldots, x_n$ be a sample of length of time that a customer is on the phone with a call center help line. We feel comfortable modeling the data as coming from an *exponential distribution*. What is the MLE (Maximum Likelihood Estimate) of the parameter? Show your steps.

::: {.callout-note title="Solution"}
Add solution here
:::


# Problem 2: Coding Practice

## a. Simulation

Simulate 100 observations from the following model:

- $X \sim N(\mu = 1, \sigma = 1)$
- $Y \sim N(\mu = 1 + 2X, \sigma = 2)$
- $Z = \begin{cases} 1 &\quad Y<0 \\ 2 &\quad Y \ge 0 \end{cases}$

::: {.callout-note title="Solution"}
Add solution here
:::


## b. Scatterplot

Make a scatter plot of the data. Put $X$ on the x-axis and $Y$ on the y-axis. Color the points according to $Z$.

::: {.callout-note title="Solution"}
Add solution here
:::


## c. Function

Write a function that adds two numbers together and squares the result.

::: {.callout-note title="Solution"}
Add solution here
:::




# Problem 3: Statistics

## a. Quantiles

Find two [quantiles](https://en.wikipedia.org/wiki/Quantile) that capture 95% of the following data:

::: {.panel-tabset}

### R

```{r}
set.seed(2026)
x = runif(n=100, min=2, max=22)
```

### Python

```{python}
import numpy as np

rng = np.random.default_rng(2026)
x = rng.uniform(low=2, high=22, size=100)
```

:::


::: {.callout-note title="Solution"}
Add solution here
:::


## b. Confidence Interval

A new machine learning model, developed by UVA researchers, uses biopsy images to predict if a child has enteropathy or celiac disease. [In a study of 102 patients](https://jamanetwork.com/journals/jamanetworkopen/fullarticle/2735765), the model was able to correctly classify 95 of the images.

Find the 90% [confidence interval](https://en.wikipedia.org/wiki/Confidence_interval) for the probability a patient is correctly classified?

::: {.callout-note title="Solution"}
Add solution here
:::


## c. Linear Models

- Albemarle County, VA real estate assessment data can be found [at this link](https://raw.githubusercontent.com/uvastatlab/phdplus/master/data/albemarle_homes.csv). Data collected by UVAs StatLab as part of the PhD plus program (https://github.com/uvastatlab/phdplus). 

1. Fit a linear regression model that predicts the `totalvalue` using the predictors: `condition`, size (`finsqft`), and location (`city`).

2. What are the fitted model parameters (i.e., the estimated coefficients)?

3. What is the estimated `totalvalue` for home with the following characteristics?


| finsqft | city        | condition |
| ------: | ----------- | --------- |
|    2500 | EARLYSVILLE | Good      |
|    1850 | CROZET      | Fair      |


::: {.callout-note title="Solution"}
Add solution here
:::


## d. Hypothesis Testing

Use the `movies` (IMDb) data from the [ModernDive book](https://moderndive.com/v2/) to performance a hypothesis test that *Action* movies are ranked lower (on average) than *Romance* movies.

[movies.csv](https://raw.githubusercontent.com/moderndive/ModernDive_book/refs/heads/v2/data/movies.csv)

::: {.callout-note title="Solution"}
Add solution here
:::
